Spatiotemporal Multi-Resolution Approximation of the Amari Type Neural Field Model

### 2012-08-10 Multiresolution_framework/ltx/response_to_reviews.tex

				\item Similar claims to the ones mentioned here regarding  implications for ``patient-specific'' modelling have also been made in Freestone et al., so it is not clear how these relate to the current multi-resolution approach and also such remarks might look too ambitious given that only synthetic data with fixed parameters are considered.

				\emph{Response here.}

				\item The comments about model complexity and parsimony on p.13 are somehow vague as these quantities are undefined.

				% \emph{\parham{some notes to be added to the manuscript: Parsimony: as few as necessary variables to describe a phenomenon, complexity: to add detail description to the model to increase its accuracy with the natural phenomenon, finding a balance between complexity and parsimony is crucial. By parsimony we sacrifice realism ...  }}

				\emph{\ken{To clarify this comment, a short paragraph has been include in section 6 to highlight the modelling choice facing every user, that is; choosing between a more parsimonious or more complex model structure.}}

### 2012-08-11 Multiresolution_framework/ltx/response_to_reviews.tex

        \item Since the methodology used is the same as in that earlier Neuroimage paper, the paper at hand repeats several of  the results of that previous paper, see sections 3 and 4: equations (2)-(5) and (20) -(33) have already appeared or are very similar to equations in Freestone et al.

	\emph{Response here.}

        \item Section 5 is a review of known results and standard background material from wavelet theory and includes a simple calculation of the Fourier transform of the particular basis used here, which would be more appropriate for an appendix (cf appendix H of the Freestone et al. paper where the Fourier transform of the Gaussian basis functions is calculated).

		\emph{\parham{Section 5 summarizes important characteristics of B-Spline scaling and wavelet functions which are collected from different references, with an attempt to clarify the inner product and convolution of theses functions (please see R2 comment on ``B-spline calculus''). We think that this brief review can help the interested readers to get familiar with the B-spline scaling and wavelet functions and direct them to the important literature on this matter. As the reviewer suggested the calculation of the Fourier transform is now moved to the appendix.}}

### 2012-08-11 Multiresolution_framework/ltx/response_to_reviews.tex

        \item Section 5 is a review of known results and standard background material from wavelet theory and includes a simple calculation of the Fourier transform of the particular basis used here, which would be more appropriate for an appendix (cf appendix H of the Freestone et al. paper where the Fourier transform of the Gaussian basis functions is calculated).

		\emph{\parham{Section 5 summarizes important characteristics of B-Spline scaling and wavelet functions which are collected from different references, with an attempt to clarify the inner product and convolution of theses functions (please see R2 comment on ``B-spline calculus''). We think that this brief review can help the interested readers to get familiar with the B-spline scaling and wavelet functions and direct them to the important literature on this matter. As the reviewer suggested the calculation of the Fourier transform is now moved to the appendix.}}

        \item Section 6 also borrows heavily from Dewar et al, e.g. cf equation (58) of this paper with equation (19) of Dewar et al: the state transition matrix is a perturbation of the corresponding matrix used in that earlier paper. The approach used in both papers seems to be the same with an RTS smoother used in the E-step (this was also used and explained in Freestone et al.) and an M step optimizing a cost function, quadratic in the parameters, cf. equation (62) with  equation (29) of Theorem 2 of Dewar et al. In general, the derivations of section 6 and appendix B follow closely or are very similar to those of Dewar et al.

		\emph{\parham{As the reviewer correctly pointed out there are similarities in the derivations herein and that of Dewar et al. Our main aim here is to adapt techniques from control systems theory and machine learning, modify and apply them to the Amari type neural field models. As the result some of the derivations are somehow similar, however, there are important changes in the framework highlighted as follows. The inverse framework in Dewar et al. is based on a general IDE model which is not suitable when electrophysiological data is considered.  The inclusion of the membrane time constant and the firing rate in this work modify the general IDE equations to account for this type of data. As the result of these necessary modifications  the membrane parameter $\xi=1-T_s/\tau$ and the firing rate slope $\varsigma$ appear in the transition matrix. The novel implementation of the M-step in this work significantly simplifies the computational complexity of the EM based algorithm proposed by Dewar et al. as it is shown in Table.~4 of the manuscript, namely from $O(n_x^4n_{\theta}^2)$ to $O(n_x^2n_{\theta}^2)$, see equation (64) of this paper and equation (31) of Dewar et al, where the static and dynamic components are separated by defining $\boldsymbol \Gamma_1$ and $\mathbf U$ (see equations (65) and (67)). Therefore only the dynamic elements are updated at each iteration  to calculate $\boldsymbol\Upsilon$. The same simplification is also performed to calculate $\boldsymbol\upsilon_0, \boldsymbol\upsilon_1$ (see equations (62) and (63) of this paper and equation (30) of Dewar et al.). These modifications are significant and crucial for high dimensional systems and more specifically for implementing the multi-resolution framework introduced here which otherwise would be possible. The efficient M-step provided herein is also applicable in the framework of Freestone et al. if only the linear regime of the firing rate is considered.  The use of the RTS smoother in the E-step of the EM algorithm for state-space models is an effective, standard and widely used approach. In fact, the aim of the IDE representation in a state-space form is to adapt the EM and consequently the RTS smoother in the E-step. The RTS algorithm is presented only for completeness and it is different than Unscented RTS (for non-linear state-space models) in Freestone et al. The cost function of a linear state-space model with separable parameters is always quadratic in parameters in the M-step and in fact it is one of the advantages of this representation, giving a closed form solution for system's parameters which is essentially equivalent to minimum least squares error.}}

@@

        \item Also, Appendix A is the analogue of Appendix D of Freestone et al. and Appendix C includes results known from the literature (as the authors point out).

\emph{\parham{In order to formulate the state-space representation of the multi-resolution neural field model the inner product of the B-spline functions are required. We believe that it is necessary to provide the interested reader with the required tools to implement the algorithm. Appendix D of Freestone et al. and Appendix A herein are only similar in the sense that the inner product of two basis functions are calculated. In Appendix D of Freestone et al. two  n-dimensional Gaussians are considered while here the inner product of two B-spline functions is calculated. As for Appendix C, although authors have no objections to remove the presented derivations, we think that it would be informative for the broad audience of NeuroImage to include the derivations of some of the well-known formulations in the EM algorithm.} }

@@

				\emph{Response here.}

				\item The comments about model complexity and parsimony on p.13 are somehow vague as these quantities are undefined.

				% \emph{\parham{some notes to be added to the manuscript: Parsimony: as few as necessary variables to describe a phenomenon, complexity: to add detail description to the model to increase its accuracy with the natural phenomenon, finding a balance between complexity and parsimony is crucial. By parsimony we sacrifice realism ...  }}

				\emph{\ken{To clarify this comment, a short paragraph has been included in section 6 to highlight the modeling choice facing every user, that is; choosing between a more parsimonious or more complex model structure.}}

### 2012-08-11 Multiresolution_framework/ltx/manuscript.tex
1.11

Additional extension for future work include extending the model to capture a second order synaptic response kernel. The second-order postsynaptic kernel proposed in \citet{VanRotterdam1982} leads to dynamics that are more reminiscent to real postsynaptic potentials ( see \citet{Spiegler2010,Spiegler2011} for example). Another extension of this framework is the inclusion of finite distant-dependent action potential propagation velocity into the parametric neural field equations. In small networks the resulting delays via propagation can be neglected. However, delays must be accounted for in large scale networks. For instance, \citet{Ghosh2008} demonstrated that inclusion of time delays arising from propagation along connecting fibres are essential for the understanding of the fluctuations in resting state networks (see \citet{Deco2010} for review). In fact, \citet{Ghosh2008} emphasize that the couplings in large-scale brain networks have a space-time structure composed of their anatomical connectivity (space) and signal transmission delays (time). To capture correctly the constraints posed upon the network dynamics by connectivity, the space-time structure of the brain connectivity must be correctly estimated. Finally, future work should be directed towards applying and validating the framework on real data. 

>>>>

Additional extension for future work include extending the model to capture a second order synaptic response kernel. The second-order postsynaptic kernel proposed in \citet{VanRotterdam1982} leads to dynamics that are more reminiscent to real postsynaptic potentials ( see \citet{Spiegler2010,Spiegler2011} for example). Another extension of this framework is the inclusion of finite distant-dependent action potential propagation velocity into the parametric neural field equations. In small networks the resulting delays via propagation can be neglected. However, delays must be accounted for in large scale networks. For instance, \citet{Ghosh2008} demonstrated that inclusion of time delays arising from propagation along connecting fibres are essential for the understanding of the fluctuations in resting state networks. Finally, future work should be directed towards applying and validating the framework on real data.

% (see \citet{Deco2010} for review) In fact, \citet{Ghosh2008} emphasize that the couplings in large-scale brain networks have a space-time structure composed of their anatomical connectivity (space) and signal transmission delays (time). To capture correctly the constraints posed upon the network dynamics by connectivity, the space-time structure of the brain connectivity must be correctly estimated. 
